---
title: "5-model-summary"
author: "Daniel Hocking"
date: "August 28, 2014"
output: html_document
---

```{r}
rm(list = ls()) 
gc(verbose = FALSE)

library(ggplot2)
library(ggmcmc)
library(dplyr)
#library(nlme)
library(devtools)
#install_github("Conte-Ecology/conteStreamTemperature")
library(conteStreamTemperature)
library(rjags)
# need to load data.fixed, data.random.sites, data.random.years

baseDir <- getwd() # works as long as you have the project loaded in RStudio - does not work for kniting

dataInDir <- paste0(baseDir, '/dataIn/')
dataOutDir <- paste0(baseDir, '/dataOut/')
dataLocalDir <- paste0(baseDir, '/localData/')
graphsDir <- paste0(baseDir, '/graphs/')

load('localData/jags.RData')
load(paste0(dataLocalDir, 'tempDataSync.RData'))
load(paste0(dataLocalDir, "covariate_list.RData"))

# memory.limit(size = 1e6)

#system.time(summary.stats <- summary(M.ar1)) # takes a long time for full dataset. look for better solution


```


## Evaluate MCMC Iterations

```{r check model convergence and mixing}
system.time(ggs.ar1 <- ggs(M.ar1)) # slow and memory intensive - 2 minutes with residuals saved in mcmc object and no other programs running but uses most of the memory on my 8 GB laptop
gc(verbose = FALSE) # clears all of the temporary variables stored by ggs and cuts hanging memory more than in half.

ggmcmc(ggs.ar1, file = 'localData/ggmcmc-B0.pdf', family = "B.0", plot = c("ggs_traceplot", "ggs_compare_partial", "ggs_autocorrelation"))

ggmcmc(ggs.ar1, file = 'localData/ggmcmc-mu-huc.pdf', family = "mu.huc", plot = c("ggs_traceplot", "ggs_compare_partial", "ggs_autocorrelation"))

ggmcmc(ggs.ar1, file = 'localData/ggmcmc-mu-year.pdf', family = "mu.year", plot = c("ggs_traceplot", "ggs_compare_partial", "ggs_autocorrelation"))

ggmcmc(ggs.ar1, file = 'localData/ggmcmc-mu-ar1.pdf', family = "mu.ar1", plot = c("ggs_traceplot", "ggs_compare_partial", "ggs_autocorrelation"))

ggmcmc(ggs.ar1, file = 'localData/ggmcmc-sigma-ar1.pdf', family = "sigma.ar1", plot = c("ggs_traceplot", "ggs_compare_partial", "ggs_autocorrelation"))

ggmcmc(ggs.ar1, file = 'localData/ggmcmc-sigma-site.pdf', family = "sigma.b.site", plot = c("ggs_traceplot", "ggs_compare_partial", "ggs_autocorrelation"))

ggmcmc(ggs.ar1, file = 'localData/ggmcmc-sigma-huc.pdf', family = "sigma.b.huc", plot = c("ggs_traceplot", "ggs_compare_partial", "ggs_autocorrelation"))

ggmcmc(ggs.ar1, file = 'localData/ggmcmc-sigma-year.pdf', family = "sigma.b.year", plot = c("ggs_traceplot", "ggs_compare_partial", "ggs_autocorrelation"))

ggmcmc(ggs.ar1, file = 'localData/ggmcmc-ar1-rho-huc.pdf', family = "rho.B.huc", plot = "ggs_traceplot")

ggmcmc(ggs.ar1, file = 'localData/ggmcmc-ar1-B-ar1.pdf', family = "B.ar1", plot = c("ggs_traceplot", "ggs_compare_partial", "ggs_autocorrelation"))

gc(verbose = FALSE)

```

```{r means tables with rHat}

intersectSeveral <- function(...) { Reduce(intersect, list(...)) } 

# exclude the variables in the grep() statement
# intersect works on pairs. the intersectSeveral function generates the intersection of multiple lists
keepRows <- 
  intersectSeveral(grep("B.site", x=ggs.ar1$Parameter,invert=TRUE),
                   grep("B.huc", x=ggs.ar1$Parameter,invert=TRUE),
                   grep("B.year", x=ggs.ar1$Parameter,invert=TRUE),
                   grep("B.ar1", x=ggs.ar1$Parameter,invert=TRUE),
                   grep("rho.B.huc", x=ggs.ar1$Parameter,invert=TRUE),
                   grep("rho.B.year",x=ggs.ar1$Parameter,invert=TRUE)
                  )

rHat <- unique( ggs_Rhat(ggs.ar1[keepRows,])$data )

ggmcmc(ggs.ar1[keepRows,], file = "localData/ggmcmc-ar1-Rhat", plot = c("ggs_caterpillar"))

```

## Check model fit

```{r check model fit}
#------ Check residual patterns------

system.time(coef.summary <- avgCoefs(ggs.ar1)) # 15s
#tempDataSyncS$resid.ar1 <- dplyr::filter(coef.summary, grepl('^residuals', coef.summary$Parameter))$mean

gc(verbose = FALSE)

B.fixed <- dplyr::filter(coef.summary, grepl('^B.0', coef.summary$Parameter))
B.fixed$coef <- cov.list$fixed.ef

mu.huc <- nameCoefs(coef.summary = coef.summary, rand.levels = levels(as.factor(tempDataSyncS$HUC8)), family = "mu.huc", name = "coef", conditional = FALSE, coefs = cov.list$site.ef)

mu.year <- nameCoefs(coef.summary = coef.summary, rand.levels = levels(as.factor(tempDataSyncS$year)), family = "mu.year", name = "coef", conditional = FALSE, coefs = cov.list$year.ef)

mu.ar1 <- dplyr::filter(coef.summary, grepl('^mu.ar1', coef.summary$Parameter))
mu.ar1$coef <- mu.ar1$Parameter

sigma.b.site <- nameCoefs(coef.summary = coef.summary, rand.levels = levels(as.factor(tempDataSyncS$site)), family = "sigma.b.site", conditional = FALSE, coefs = cov.list$site.ef, name = "coef")

sigma.b.huc <- nameCoefs(coef.summary = coef.summary, rand.levels = levels(as.factor(tempDataSyncS$site)), family = "sigma.b.huc", conditional = FALSE, coefs = cov.list$site.ef, name = "coef")

sigma.b.year <- nameCoefs(coef.summary = coef.summary, rand.levels = levels(as.factor(tempDataSyncS$year)), family = "sigma.b.year", conditional = FALSE, coefs = cov.list$year.ef, name = "coef")

sigma.ar1 <- dplyr::filter(coef.summary, grepl('^sigma.ar1', coef.summary$Parameter))

B.ar1 <- nameCoefs(coef.summary = coef.summary, rand.levels = levels(as.factor(tempDataSyncS$site)), family = 'B.ar1', name = "site")

B.site <- nameCoefs(coef.summary = coef.summary, rand.levels = levels(as.factor(tempDataSyncS$site)), family = "B.site", coefs = cov.list$site.ef, name = "site")

B.huc <- nameCoefs(coef.summary = coef.summary, rand.levels = levels(as.factor(tempDataSyncS$HUC8)), family = "B.huc", coefs = cov.list$site.ef, name = "huc")

B.year <- nameCoefs(coef.summary = coef.summary, rand.levels = levels(as.factor(tempDataSyncS$year)), family = "B.year", coefs = cov.list$year.ef, name = "year")


# Make correlation matrix of random huc effects
cor.huc <- as.data.frame(matrix(NA, length(cov.list$site.ef), length(cov.list$site.ef)))
names(cor.huc) <- cov.list$site.ef
row.names(cor.huc) <- cov.list$site.ef
for(k in 1:length(cov.list$site.ef)){
  for(k.prime in 1:length(cov.list$site.ef)){
    cor.huc[k, k.prime] <- coef.summary[which(coef.summary$Parameter == paste('rho.B.huc[',k,',',k.prime,']', sep="")), "mean"]
  }
}
cor.huc <- round(cor.huc, digits=3)
cor.huc[upper.tri(cor.huc, diag=TRUE)] <- ''
cor.huc

# Make correlation matrix of random year effects
cor.year <- as.data.frame(matrix(NA, length(cov.list$year.ef), length(cov.list$year.ef)))
names(cor.year) <- cov.list$year.ef
row.names(cor.year) <- cov.list$year.ef
for(l in 1:length(cov.list$year.ef)){
  for(l.prime in 1:length(cov.list$year.ef)){
    cor.year[l, l.prime] <- coef.summary[which(coef.summary$Parameter == paste('rho.B.year[',l,',',l.prime,']', sep="")), "mean"]
  }
}
cor.year <- round(cor.year, digits=3)
cor.year[upper.tri(cor.year, diag=TRUE)] <- ''
cor.year


fix.ef <- rbind(B.fixed, select(mu.huc, -index), select(mu.year, -index), mu.ar1)
rand.ef <- rbind(select(sigma.b.site, -index), select(sigma.b.huc, -index), select(sigma.b.year, -index), mutate(sigma.ar1, coef = Parameter))

mod.summary <- rbind(fix.ef, rand.ef)

# table of parameter estimates for MS
mod.summary <- left_join(mod.summary, rHat[,c('Parameter','Rhat')], by = "Parameter")


coef.list <- list(fix.ef = fix.ef 
                  ,rand.ef = rand.ef
                  ,B.fixed = B.fixed
                  , mu.huc = mu.huc
                  , mu.year = mu.year
                  , mu.ar1 = mu.ar1
                  , sigma.site = sigma.b.site
                  , sigma.huc = sigma.b.huc
                  , sigma.year = sigma.b.year
                  , sigma.ar1 = sigma.ar1
                  , cor.huc = cor.huc
                  , cor.year = cor.year
                  , B.site = B.site
                  , B.huc = B.huc
                  , B.year = B.year
                  , B.ar1 = B.ar1
                  )

save(coef.list, file = paste0(dataLocalDir, "coef.Rdata"))
write.table(mod.summary, file = paste0(dataLocalDir, "modle_summary.csv"), sep = ",", row.names = F)

```






```{r figures}
# Effects of Forest
# Plot effect of catchment forest on occurrence prob at a typical HUC10 basin # Gelman p. 44
eff.forest <- data.frame(Forest=seq(0,100,length.out=100))
eff.forest$forest <- as.numeric(stdCovs(eff.forest, tempDataSync, varNames=c("Forest"))$Forest)

fixEf <- modSummary@fixEf$Mean
names(fixEf) <- row.names(modSummary@fixEf)
UCI <- modSummary@fixEf$UCI
names(UCI) <- row.names(modSummary@fixEf)
LCI <- modSummary@fixEf$LCI
names(LCI) <- row.names(modSummary@fixEf)

eff.forest$mean <- fixEf["intercept"] + fixEf["forest"]*eff.forest$forest
eff.forest$lower <- fixEf["intercept"] + LCI["forest"]*eff.forest$forest
eff.forest$upper <- fixEf["intercept"] + UCI["forest"]*eff.forest$forest

ggplot(eff.forest, aes(x = Forest, y = mean)) + 
  geom_ribbon(data=eff.forest, aes(ymin = lower, ymax = upper), fill="grey") +
  geom_line(colour = "black", size = 1) +
  xlab("Percent forest cover upstream") +
  ylab("Stream temperature (C)") +
  theme_bw() + 
  ylim(15, 25) +
  theme(axis.text.y = element_text(size=15),
        axis.text.x = element_text(size=15),
        axis.title.x = element_text(size=17, face="bold"),
        axis.title.y = element_text(size=17, angle=90, face="bold"),
        plot.title = element_text(size=20))

```

Problem with above is that confidence intervals don't really represent CI for predicted temperature. Could bootstrap with the summary stats I have but that loses some of the value of MCMC iterations.

Try to organize so that predictions can be done for each iteration. This could get big because it would be a prediction for each forest cover value for each iteration (~100 x ~1000-10000 depending how many iterations are saved). This would be done for each variable we want to show the effects of and each scenario we want to do predictions for.

```{r}
mat.ar1 <- unlist(as.matrix(M.ar1))

# how to replace or keep track of the coefficient names (e.g. "B.0[1]", "B.Site[1,1]")
n.iter <- dim(mat.ar1)[1]
fix.eff <- data.frame(matrix(NA, n.iter, K.0 + K + L))
names(fix.eff) <- c(names(data.fixed), names(data.random.sites), names(data.random.years))
  for(k in 1:K.0){
    fix.eff[ , k] <- mat.ar1[ , paste0('B.0[',k,']')]
  }
  for(k in 1:K){
    fix.eff[ , k+K.0] <- mat.ar1[ , paste0('mu.huc[',k,']')]
  }
  for(l in 1:L){
    fix.eff[ , l+K.0+K] <- mat.ar1[ , paste0('mu.year[',l,']')]
  }
str(fix.eff)
head(fix.eff)

```

## the below works so now turn into a function

```{r figures}
# Effects of Forest
# Plot effect of catchment forest on occurrence prob at an average site for a typical HUC8 basin # Gelman p. 44
eff.forest <- data.frame(Forest=seq(0,100,length.out=100))
eff.forest$forest <- as.numeric(stdCovs(eff.forest, tempDataSync, varNames=c("Forest"))$Forest)

temp.forest <- matrix(NA, n.iter, dim(eff.forest)[1])
for(i in 1:n.iter) {
  temp.forest[i, ] <- fix.eff$intercept[i] + fix.eff$forest[i]*eff.forest$forest
  }
eff.forest$mean <- apply(temp.forest, MARGIN = 2, mean)
eff.forest$lower <- apply(temp.forest, MARGIN = 2, quantile, probs = 0.025)
eff.forest$upper <- apply(temp.forest, MARGIN = 2, quantile, probs = 0.975)

ggplot(eff.forest, aes(x = Forest, y = mean)) + 
  geom_ribbon(data=eff.forest, aes(ymin = lower, ymax = upper), fill="grey") +
  geom_line(colour = "black", size = 1) +
  xlab("Percent forest cover upstream") +
  ylab("Stream temperature (C)") +
  theme_bw() + 
  ylim(15, 25) +
  theme(axis.text.y = element_text(size=15),
        axis.text.x = element_text(size=15),
        axis.title.x = element_text(size=17, face="bold"),
        axis.title.y = element_text(size=17, angle=90, face="bold"),
        plot.title = element_text(size=20))

```

```{r figures}
load("dataOut/tempDataSync-northeast.Rdata")
load("dataOut/coef-northeast.Rdata")
# Effects of Impoundments
# Plot effect of catchment forest on occurrence prob at an average site for a typical HUC8 basin # Gelman p. 44
eff.impound <- data.frame(ImpoundmentsAllSqKM = seq(min(tempDataSync$ImpoundmentsAllSqKM), 200, length.out=100))
eff.impound$ImpoundmentsAllSqKM.s <- as.numeric(stdCovs(eff.impound, tempDataSync, var.names=c("ImpoundmentsAllSqKM"))$ImpoundmentsAllSqKM)

stdCovs <- function(x, y, var.names){
  for(i in 1:length(var.names)){
    x[ , var.names[i]] <- (x[ , var.names[i]] - mean(y[[var.names[i]]], na.rm = T)) / sd(y[[var.names[i]]], na.rm = T)
  }
  return(x)
}

eff.impound$mean <- NA
eff.impound$lower <- NA
eff.impound$upper <- NA
for(i in 1:length(eff.impound$ImpoundmentsAllSqKM)){
eff.impound$mean[i] <- as.numeric(fix.ef[which(fix.ef$coef == "intercept"), "mean"] + fix.ef[which(fix.ef$coef == "ImpoundmentsAllSqKM"), "mean"] * eff.impound$ImpoundmentsAllSqKM.s[i])
eff.impound$lower[i]  <- as.numeric(fix.ef[which(fix.ef$coef == "intercept"), "mean"] + fix.ef[which(fix.ef$coef == "ImpoundmentsAllSqKM"), "qLo"] * eff.impound$ImpoundmentsAllSqKM.s[i])
eff.impound$upper[i]  <- as.numeric(fix.ef[which(fix.ef$coef == "intercept"), "mean"] + fix.ef[which(fix.ef$coef == "ImpoundmentsAllSqKM"), "qHi"] * eff.impound$ImpoundmentsAllSqKM.s[i])
}

ggplot(eff.impound, aes(x = ImpoundmentsAllSqKM, y = mean)) + 
  geom_ribbon(data=eff.impound, aes(ymin = lower, ymax = upper), fill="grey") +
  geom_line(colour = "black", size = 1) +
  xlab("Upstream Impounded Area (sq km)") +
  ylab("Stream temperature (C)") +
  theme_bw() + 
  ylim(15, 25) +
  theme(axis.text.y = element_text(size=15),
        axis.text.x = element_text(size=15),
        axis.title.x = element_text(size=17, face="bold"),
        axis.title.y = element_text(size=17, angle=90, face="bold"),
        plot.title = element_text(size=20))



```

```{r figures}
# Effects of Forest
# Plot effect of catchment forest on occurrence prob at an average site for a typical HUC8 basin # Gelman p. 44
eff.forest <- data.frame(airTemp=seq(-5, 30, length.out=100))
eff.forest$airTempS <- as.numeric(stdCovs(eff.forest, tempDataSync, varNames=c("airTemp"))$airTemp)

temp.forest <- matrix(NA, n.iter, dim(eff.forest)[1])
for(i in 1:n.iter) {
  temp.forest[i, ] <- fix.eff$intercept[i] + fix.eff$airTemp[i]*eff.forest$airTempS
  }
eff.forest$mean <- apply(temp.forest, MARGIN = 2, mean)
eff.forest$lower <- apply(temp.forest, MARGIN = 2, quantile, probs = 0.025)
eff.forest$upper <- apply(temp.forest, MARGIN = 2, quantile, probs = 0.975)

ggplot(eff.forest, aes(x = airTemp, y = mean)) + 
  geom_ribbon(data=eff.forest, aes(ymin = lower, ymax = upper), fill="grey") +
  geom_line(colour = "black", size = 1) +
  xlab("Percent forest cover upstream") +
  ylab("Stream temperature (C)") +
  theme_bw() + 
  ylim(0, 25) +
  theme(axis.text.y = element_text(size=15),
        axis.text.x = element_text(size=15),
        axis.title.x = element_text(size=17, face="bold"),
        axis.title.y = element_text(size=17, angle=90, face="bold"),
        plot.title = element_text(size=20))

```



