---
title: "Predictions"
output: html_document
---

```{r load, echo=FALSE, results='hide'}
# library(rjson)
library(jsonlite)
library(ggplot2)
library(reshape2)
library(dplyr)
library(tidyr)
library(zoo)
library(lubridate)
library(DataCombine) # for the slide function
library(RPostgreSQL)
library(devtools)
#install_github("Conte-Ecology/conteStreamTemperature")
library(conteStreamTemperature)

data_dir <- "localData_2015-07-09" 
  
  baseDir <- getwd()
  
  dataInDir <- paste0(baseDir, '/dataIn/')
  dataOutDir <- paste0(baseDir, '/dataOut/')
  dataLocalDir <- paste0(baseDir, '/localData/')
  graphsDir <- paste0(baseDir, '/graphs/')
  
  #source(paste0(baseDir, 'code/functions/temperatureModelingFunctions.R'))
  #source(paste0(baseDir, 'code/functions/dataIndexingFunctions.R'))
  
  #load((paste0(dataOutDir, 'modSummary.RData')))
coef.list <- readRDS(paste0(dataLocalDir, 'coef.RData'))
  load(paste0(dataLocalDir, 'tempDataSync.RData'))
  #load(paste0(dataLocalDir, 'covariate_list.RData'))
  #load(paste0(dataLocalDir, 'springFallBreakpoints.RData'))

cov.list <- readRDS(paste0(dataLocalDir, 'covariate-list.RData'))
springFallBPs <- readRDS(paste0(dataLocalDir, 'springFallBPs.RData'))
  
```
  
  

## Statistical Model Background

Statistical models of stream temperature often rely on the close relationship between air temperature and water temperature. However, this relationship breaks down during the winter in temperature zones, particularly as streams freeze, thereby changing their thermal and properties. Many researchers and managers are interested in the non-winter effects of temperature. The winter period, when phase change and ice cover alter the air-water relationship, differs in both time (annually) and space. We developed an index of air-water synchrony ($Index_{sync}$) so we can model the portion of the year that it not affected by freezing properties. The index is the difference between air and observed water temperatures divided by the water temperature plus 0.000001 to avoid division by zero. 

We calculate the $Index_{sync}$ for each day of the year at each reach for each year with observed data. We then calculate the 99.9% confidence interval of $Index_{sync}$ for days between the 125 and 275 days of the year (05 May and 02 October). Then moving from the middle of the year (day 180) to the beinging of the year, we searched for the first time when 10 consecutive days were not within the 99.9% CI. This was selected as the spring breakpoint. Similarly moving from the middle to the end of the year, the first event with fewer than 16 consecutive days within the 99.9% CI was assigned as the autumn breakpoint. Independent breakpoints were estimated for each reach-year combination. For reach-years with insufficient data to generate continuous trends and confidence intervals, we used the mean break points across years for that reach. If there was not sufficient local reach information, we used the mean breakpoints from the smallest hydrologic unit the reach is nested in (i.e. check for mean from HUC12, then HUC10, HUC8, etc.). More details regarding the identification of the synchronized period can be found in Letcher et al. (*in review*). The portion of the year between the spring and autumn breakpoints was used for modeling the non-winter, approximately ice-free stream temperatures.

We used a generalized linear mixed model to account for correlation in space (stream reach nested within HUC8). This allowed us to incorporate short time series as well as long time series from different reaches and disjunct time series from the same reaches without risk of pseudoreplication (ref: Hurlbert). By limited stream drainage area to <400 $km^2$ and only modeling the synchronized period of the year, we were able to use a linear model, avoiding the non-linearities that occur at very high temperatures due to evaporative cooling and near 0 C dur to phase change (ref: mohseni). 

We assumed stream temperature measurements were normally distributed following,

$$ t_{h,r,y,d} \sim \mathcal{N}(\mu_{h,r,y,d}, \sigma) $$

where $t_{h,r,y,d}$ is the observed stream water temperature at the reach ($r$) within the sub-basin identified by the 8-digit Hydrologic Unit Code (HUC8; $h$) for each day ($d$) in each year ($y$). We describe the normal distribution based on the mean ($mu_{h,r,y,d}$) and standard deviation ($\sigma$) and assign a vague prior of $\sigma = 100$. The mean temperature is modelled to follow a linear trend

$$ \omega_{h,r,y,d} = X_0 B_0 + X_{h,r} B_{h,r} + X_{h} B_{h} + X_{y} B_{y} $$

but the expected mean temperature ($\mu_{h,r,y,d}$) was also adjusted based on the residual error from the previous day

$$ \mu_{h,r,y,d} = \begin{cases}
    \omega_{h,r,y,d} + \delta_{r}(t_{h,r,y,d-1} - \omega_{h,r,y,d-1}) & \quad  \text{for $t_{h,r,y,d-1}$ is real} \\
    \omega_{h,r,y,d} & \quad  \text{for $t_{h,r,y,d-1}$ is not real}
  \end{cases}
 $$

where $\delta_{r}$ is an autoregressive [AR(1)] coefficient that varies randomly by reach and $\omega_{h,r,y,d}$ is the expected temperature before accounting for temporal autocorrelation in the error structure.

$X_{0}$ is the $n \times K_0$ matrix of predictor values. $B_0$ is the vector of $K_0$ coefficients, where $K_0$ is the number of fixed effects parameters including the overall intercept. We used **???XX???** fixed effect parameters including the overall intercept. These include **??latitude, longitude, upstream drainage area, percent forest cover, elevation, surficial coarseness classification, percent wetland area, upstream impounded area, and an interaction of drainage area and air temperature??**. We assumed the following distributions and vague priors for the fixed effects coefficients

$$ B_0 \sim \mathcal{N}(0,\sigma_{k_0}), \text{for $k_0 = 1,...,K_0$,} $$

$$ B_0 = \beta_{0}^{1},...,\beta_{0}^{K_{0}} \sim \mathcal{N}(0, 100) $$

$$ \sigma_{k_0} = 100 $$

**??The effects of air temperature on the day of observation ($d$) and mean air temperature over the previous 7 days varied randomly with reach nested within HUC8, as did precipitation, the previous 30-day precipitation mean, and the interactions of air temperature and preciptation (all 4 combinations).??**

$B_{h,r}$ is the $R \times K_{R}$ matrix of regression coefficients where $R$ is the number of unique reaches and $K_{R}$ is the number of regression coeffcients that vary randomly by reach within HUC8. We assumed prior distributions of

$$ B_{h,r} \sim \mathcal{N}(0,\sigma_{k_{r}}), \text{for $k_{r} = 1,...,K_{R}$,} $$

$$ \sigma_{r_0} = 100 $$

$X_{h}$ is the matrix of parameters that vary by HUC8. We allowed for correlation among the effects of these HUC8 coefficients as described by Gelman and Hill [-@Gelman2007].

$B_{h}$ is the $H \times K_{H}$ matrix of coefficients where $H$ is the number of HUC8 groups and $K_H$ is the number of paramaters that vary by HUC8 including a constant term. In our model, $K_{H} = K_{R}$ and we assumed priors distributions of

$$ B_{h} \sim \mathcal{N}(M_{h},\Sigma_{B_{h}}), \text{for $h = 1,...,H$} $$

where $M_{h}$ is a vector of length $K_{H}$ and $\Sigma_{B_{h}}$ is the $K_{H} \times K_{H}$ covariance matrix.

$$ M_{h} \sim MVN(\mu_{1:K_h}^h, \sigma_{1:K_h}^h) $$

$$ \mu_{1}^h = 0; \mu_{2:K_h}^h \sim \mathcal{N}(0, 100) $$

$$ \Sigma_{B_{h}} \sim \text{Inv-Wishart}(diag(K_{h}), K_{h}+1) $$

Similarly, we allowed the some effects of some parameters ($X_{y}$) to vary randomly by year with potential correlation among the coefficients. The intercept, day of the year ($day$), $day^2$, and $day^3$ all varied randomly with year such that $K_{y} = 4$. We assumed prior distributions of

$$ B_{y} \sim \mathcal{N}(M_{y},\Sigma_{B_{y}}), \text{for $y = 1,...,Y$} $$

where $M_{y}$ is a vector of length $K_{Y}$ and $\Sigma_{B_{y}}$ represents the $K_{Y} \times K_{Y}$ covariance matrix.

$$ M_{y} \sim MVN(\mu_{1:K_y}^y, \sigma_{1:K_y}^y) $$

$$ \mu_{1}^y = 0; \mu_{2:K_y}^y \sim \mathcal{N}(0, 100) $$

$$ \Sigma_{B_{y}} \sim \text{Inv-Wishart}(diag(K_{y}), K_{y}+1) $$

To estimate all the parameters and their uncertainties, we used a Bayesian analysis with a Gibbs sampler implemented in JAGS (ref) through R (ref) using the rjags package (ref). This approach was beneficial for hierarchical model flexibility and tractibility for large datasets. We used vague priors for all parameters so all inferences would be based on the data.


# Predictions

### Pull Landscape & Land-use Data Associated with particular `featureid`

All spatial information in the database is centered around the `featureid`. Select the `featureid` or IDs of interest for the prediction. Then pull the associated landscape and land-use information (basin characteristics). Information on these metrics can be found in the GitHub repo: [https://github.com/Conte-Ecology/basinCharacteristics](https://github.com/Conte-Ecology/basinCharacteristics).


### Pull Temporal Data (from daymet)

### Clip to Synchronized Time Period

Example of breakpoints file:

```{r breakpoints, echo=FALSE}
str(springFallBPs)
head(springFallBPs)
```

### Pull Observed Data if Available

### Standardize Covariates (independent data)

Example of standardization file:

```{r breakpoints, echo=FALSE}
str(df_stds)
head(df_stds)
```

### Assign Conditional or Marginal Coefficient Estimates for Random Effects

Example of covariate list:

```{r breakpoints, echo=FALSE}
str(cov.list)
head(cov.list)
```

Example of coeficient data:

```{r breakpoints, echo=FALSE}
str(coef.list)
head(coef.list)
```

### Calculate Trend

```
  df$trend <- NA
  df$trend <- as.vector(coef.list$B.fixed$mean %*% t(as.matrix(as.data.frame(unclass(select(ungroup(data), one_of(cov.list$fixed.ef))))))) +
    rowSums(as.matrix(select(df, one_of(cov.list$site.ef))) * as.matrix(select(df, one_of(names(B.site[-1]))))) +
    rowSums(as.matrix(select(df, one_of(cov.list$huc.ef))) * as.matrix(select(df, one_of(names(B.huc[-1]))))) +
    rowSums(as.matrix(select(df, one_of(cov.list$year.ef))) * as.matrix(select(df, one_of(names(B.year[-1])))))
  
  # Add B.ar1 to predictions
  df <- mutate(df, prev.temp = c(NA, temp[(2:(nrow(data))) -1]))
  df <- mutate(df, prev.trend = c(NA, trend[(2:nrow(data)) - 1]))
  df <- mutate(df, prev.err = prev.temp - prev.trend)
  df <- mutate(df, tempPredicted = trend)
  df[which(!is.na(df$prev.err)), ]$tempPredicted <- df[which(!is.na(df$prev.err)), ]$trend + df[which(!is.na(df$prev.err)), ]$B.ar1 * df[which(!is.na(df$prev.err)), ]$prev.err
  
```

### Add Autoregressive Error




